---
title: "Data Exploration for DSNLI project."
output: html_notebook
---

Load needed libraries.
```{r, echo=FALSE, include=FALSE}
library(readxl)
library(plyr)  # I need some old functions from here
library(ggplot2)
library(gridExtra)
library(dplyr)
```

Custom-made functions.
```{r}
source("../r_utils/subset_on_level.r")
```


Read data.
```{r}  
postcodes = read_xls("../initial_docs/inspost.xls")
data = read.csv("../data/train.csv")
names(data)[names(data) == "X"] = "policy_id"  # so we can identify if needed

# I realized later on that we need to average the chargtot over number of claims.
data2 = data %>% 
          filter(nbrtotc > 0) %>%
          mutate(avg_claim = chargtot/nbrtotc)


set.seed(0916778)
```

These are just further info about the postal-codes.
```{r}
head(postcodes)
```

I can join this data with the latitude and longitude data if I need to. 
```{r}
head(data)
```

We do not need both exposure and log of exposure. Also, not clear what the column nbrtotan is, so I will drop it.
```{r}
data = subset(data, select=-c(lnexpo, nbrtotan))
```

Set categoricals, and merge with postcodes.
```{r}
cat_clmns = c("CODPOSS", "agecar", "sexp", "fuelc", "split", "usec",
              "fleetc", "sportc", "coverp", "powerc")
for (cat_col in cat_clmns) {
  data[,cat_col] = as.factor(data[, cat_col])
}

data = merge(data, postcodes, by="CODPOSS")
head(data)
```

See distribution of numerical data.
```{r, echo=FALSE}
par(mfrow=c(2, 2))

hist(data$AGEPH)  # right-skewed slightly
hist(data$nbrtotc)  # 1 and even 2 is not so insignificant apparently
hist(data2$avg_claim)  # hard to see anything here, will split below.
hist(data$duree)  # for the vast majority exposure is the whole year
```

We can see the Gamma distribution better. Keep in mind the different scales, and the right plot's first bucket includes zero.
```{r}
par(mfrow=c(1, 2))
hist(as.matrix(data2 %>% select(avg_claim) %>% filter(avg_claim < 20000)), main="avg_claims < 20,000")
hist(as.matrix(data2 %>% select(avg_claim) %>% filter(avg_claim >= 20000)), main="avg_claims >= 20,000")
```


The only continuous predictor we have is age, and even then I might bin it later. The plots below are not very informative, so I will plot a Poisson GLM for nrtotc vs ageph and a Gamma GLM for chargtot vs ageph real quick.
```{r, echo=FALSE}
par(mfrow=c(1, 2))
plot(cbind(data$AGEPH, data$nbrtotc), xlab="Age of Policy Holder", ylab="Number of Claims",
     main="nbrtotc vs ageph")
plot(cbind(data$AGEPH, data$chargtot), xlab="Age of Policy Holder", ylab="Claim Amount",
     main="chargtot vs ageph")
```

Quick Poisson GLM and plot. The Gamma GLM of chargtot vs ageph did not go well. We can definitely see that older people are safer drivers.
```{r, echo=FALSE}
poi_reg = glm(nbrtotc~AGEPH, data=data, family=poisson(link="log"))

# plot the resulting mean
youngest = min(data$AGEPH)
oldest = max(data$AGEPH)

ages = data.frame(seq(youngest, oldest)); colnames(ages) = c("AGEPH")
preds = predict.glm(poi_reg, ages, type="response")

plot(cbind(ages, preds), xlab="Age of policy holder", ylab="Mean Accidents",
     main="Mean nbrtotc vs AGEPH", type="l")
```

Let's see distribution of the categoricals.
```{r}
par(mfrow=c(3, 3))
factor_clmns = cat_clmns[-1]  # remove postalcodes
for (col in factor_clmns) {
  barplot(summary(data[, col]), main=col)
}
```

Relevel the factors based on the mode. I think as.factor already does this automatically but it doesn't hut to be careful.
```{r}
data = within(data, agecar<-relevel(agecar, ref="6-10"))
data = within(data, sexp<-relevel(sexp, ref="Male"))
data = within(data, fuelc<-relevel(fuelc, ref="Petrol"))
data = within(data, split<-relevel(split, ref="Once"))
data = within(data, usec<-relevel(usec, ref="Private"))
data = within(data, fleetc<-relevel(fleetc, ref="No"))
data = within(data, sportc<-relevel(sportc, ref="No"))
data = within(data, coverp<-relevel(coverp, ref="MTPL"))
data = within(data, powerc<-relevel(powerc, ref="<66"))
```


We only have one continuous predictor (that we might bin later), AGEPH. The rest of our predictors are categorical. We can do level plots (plot nbrtotc vs AGEPH given factors=(1, 0 ,...)). Let us try to select a few candidates.
```{r}
counts = plyr::count(data, vars=factor_clmns)  # intentionally ignored postal codes in this
head(counts)
print(sort(counts$freq, decreasing=TRUE)[1:10])  # size on these levels is very respectable
                                                 # keep in mind that trainset is only 64% of full data
```

Technically I could have aggregated the postcodes together as another factor, something like {Rural, Urban} or {Brussels, Antwerp, Ghent, Charleroi, Liege, others}, the most populous cities in Belgium. However, I will only use the spatial data in the GAM step.

Below are the possible settings we can use to see if there is a lack of fit for nbrtotc vs AGEPH.
```{r}
levels = counts[order(counts$freq, decreasing=TRUE), ]
head(levels, 20)
levels = subset(levels, select=-c(freq))
```

Let us see if there is a lack-of-fit between just nbrtotc vs AGEPH. Note that the points are jittered. On the left plot, we can definitely see the change of density around 0 between young (17-30ish) drivers and old (60-78ish) drivers, so the overall decreasing trend is certainly justified.

On the other hand, on the right plot we can definitely see that ideally the predictions should be higher for older drivers (60-78ish). Compare the grids between (50, 62.5] and (62.5, 70]. Keeping in mind that there are more drivers in the (50, 62.5] range, compare the densities around 0 and around 1. For the older drivers, the ratio of points (around 0) over (around 1) is smaller compared to the (50, 62.5] range. Of course, this being a single-variable GLM, we will only have a strictly monotonic prediction. This already suggests underfitting.
```{r, echo=FALSE}
set.seed(0916778)
tmp = subset_on_level(data, levels[1, ])
tmp_sample = tmp[sample(nrow(tmp), 400), ]

map1 = aes(x=AGEPH, y=nbrtotc)
map2 = aes(x=AGEPH, y=preds)
plot1 = ggplot(data=tmp_sample, mapping=map1) + geom_jitter(height=0.2, width=0, cex=0.5) + geom_line(mapping=map2, data=cbind(ages, preds))
plot1 = plot1 + labs(x="Age of policy holder", y="Number of claims", title="levels[1, ]")

set.seed(0916778)
tmp = subset_on_level(data, levels[12, ])
tmp_sample = tmp[sample(nrow(tmp), 400), ]

map1 = aes(x=AGEPH, y=nbrtotc)
map2 = aes(x=AGEPH, y=preds)
plot2 = ggplot(data=tmp_sample, mapping=map1) + geom_jitter(height=0.2, width=0, cex=0.5) + geom_line(mapping=map2, data=cbind(ages, preds))
plot2 = plot2 + labs(x="Age of policy holder", y="Number of claims", title="levels[12, ]")

grid.arrange(plot1, plot2, ncol=2)
```

Save the processed train data.
```{r}
save(data, file="../data/train.Rda")
```





